Learning from Data Through Models

### 2023-09-05 Paper/draft.Rmd
editing outline

# Framework

@@

An agent has a \emph{misspecified} belief if the prior assigns probability zero to their true type. Furthermore, the agent is 
\emph{dogmatic} if he holds a degenerate belief that places probability one on being a particular type, $\hat{\theta}$. A dogmatic 
agent can be dogmatic and misspecified. That means that $\hat{\theta} \neq \theta^*$ and $p_0(\hat{\theta}) = 1$.

The agent chooses an action $a\in A$ and observes an outcome $h\in H$. 
The outcome is a function of the agent's type, the state, and
the action. In particular $h(\theta^*, \omega^*, a)$ is increasing in both $\theta^*$ and $\omega^*$. And is such that Conditional
on a pair of parameters $(\theta, \omega)$, there is a unique optimal action.

## Example

Set $A = \Omega = $ and $H = [0, \infty)$ and consider a student with intrinsic ability $\theta^*\geq 0$ who faces a grading procedure $\omega^*$ that is unknown to them. However, 
they know that a higher $\omega^*$ is more likely to yield a higher grade. In particular, assume the grade is given by 
$(\theta^*+a)\omega^*$.

The student must choose an effort level $a$ which will
determine their grade. For whatever the chosen effort level is, the agent must pay a cost $c(a) = \frac{1}{2}a^2$. Therefore, the 
student's payoff is given by

\begin{equation}
u(a; \theta^*, \omega^*) = (\theta^*+a)\omega^* - \frac{1}{2}a^2
\end{equation}

## The Bayesian Benchmark

Bayesian agents always choose the effort level that maximizes their flow payoff by taking expectations over
their prior beliefs about $\theta$ and $\omega$. Since agents are myopic, even though all the parameters could be identified with enough variation in choices, 
a fully Bayesian agent might not learn their true type. This happens because they do not internalize the tradeoff between flow payoff
and learning and thus might not experiment enough to learn their type. An alternative to this approach is given 
by @Hestermann2021 and is discussed with the results.

## A Biased Agent
An agent that updates their beliefs with self-attribution bias will update their beliefs about the state $\omega$ and their type $\theta$
by over-attributing successes to a high value of $\theta$ and under-estimating the role of higher $\omega$. Similarly,
they will attribute failure to a low state more than an unbiased agent would. 
To model the self-serving attribution bias, I take the approach of @benjamin2019, where a generalization of the Bayes rule above gives the update.

## The Dogmatic Agent

A dogmatic agent does not update their beliefs about $\theta$; instead,
they hold a degenerate belief that places probability one on being a particular type, $\hat{\theta}$. For instance, the dogmatic agent 
might be of type $\hat{\theta} = \theta_M$ but holds a belief that places probability one on being of type $\theta_H$. In this case, no matter
how much information he gathers against being of type $\theta_H$, he will not update his beliefs. Any discrepancies between the 
observed outcomes are incorporated using the Bayes rule to update their beliefs about $\omega$. This results in the following
posterior odds:

A crucial difference between the dogmatic agent and the Bayesian agent is that the dogmatic agent does not aggregate across 
types when updating their beliefs about $\omega$. This means the dogmatic agent will never learn their true type if they are 
misspecified.

@Heidhues2018 show that in a setting such as ours, a dogmatic modeler will inevitably fall into a self-confirming equilibrium where
the outcomes they observe reinforce their belief on $\omega$ in such a way that as $t\to\infty$ the agent will be certain that
the state is some $\omega^*$ consistent with their believed type and the observed data.

EXPLAIN WHY THEY ARE ALWAYS DISAPPOINTED AND UPDATE DOWNWARDS ON OMEGA UNTIL THEY FALL INTO THE TRAP.

@@

An agent is a \emph{switcher} if they behave as a dogmatic but is willing to entertain the possibility that they are of a different type.
In particular, when they start off as a misspecified dogmatic, they are willing to switch to a different dogmatic belief if the data is
convincing enough.

@@

EXPLAIN HOW THEY MIGHT GET OUT OF THE TRAP.

# A Simple Example
The agent can be of one of 3 types: $\theta \in \{\theta_L, \theta_M, \theta_H\}$  with $\theta_H > \theta_M > \theta_L$.
They face an unknown exogenous success rate
$\omega \in \{\omega_L, \omega_M, \omega_H\}$ with $\omega_H>\omega_M>\omega_L$. Each of the values of $\omega$ is realized with equal probability. 
The agent knows the distribution of $\omega$ but not its realized value. Let the true type be $\theta^*$ 
and the true state be $\omega^*$.

@@

>>>>

# Setting 1

@@

An agent has a \emph{misspecified} belief if the prior assigns probability zero to their true type. Furthermore, the agent is 
\emph{dogmatic} if he holds a degenerate belief that places probability one on being a particular type, $\hat{\theta}$. An 
agent can be dogmatic and misspecified; that means that $\hat{\theta} \neq \theta^*$ and $p_0(\hat{\theta}) = 1$.

The agent chooses an action $a\in A$ and observes a noisy outcome $h\in H$. 
The outcome is a function of the agent's type, the state, and
the action. In particular $h = h(\theta^*, \omega^*, a) + \varepsilon$ with $h(\cdot)$ increasing in both $\theta^*$ and $\omega^*$, 
and such that conditional on a pair of parameters $(\theta, \omega)$, there is a unique optimal action. 
$\varepsilon\sim N(0, \sigma)$ is noise in the output.

After observing the outcome, the agent updates his beliefs about $\theta$ and $\omega$ using some algorithm and moves on to the next period.
He repeats this process infinitely many times. I make the simplifying assumption that the agent is myopic and chooses the action 
that maximizes the payoff in each period. This assumption is symplifies the analysis and plays a role on whether an agent who updates
 their beliefs using bayes rule would learn the truth or not.

A key notion in this setting is that of a self-defeating equilibrium\footnote{This notion is an adaptation of the Berk-Nash equilibrium in 
@Esponda2016 to this setting with only one agent}. 
A \emph{self-defeating equilibrium} is a belief and action pair such that the agent's belief about their type is misspecified and the 
outcome generated by the action is consistent with the misspecified belief. The average outcome under the true type and the true 
state equals the average output the agent expects under the misspecified belief. In addition, the agent's belief is said to be 
\emph{stable} when this happens.

Within this framework, I consider two nested theories of belief updating. The first one is a dogmatic modeler from @Heidhues2018.
The second one is a switcher, as in @Ba2023. The dogmatic modeler can be seen as a switcher with an infinitely sticky initial belief
and this is the sense in which the two are nested. Both theories produce different predictions with respect to the equilibrium outcome.

## The Dogmatic Modeler

A dogmatic agent does not update their beliefs about $\theta$; instead he holds a degenerate belief that places probability one 
on being a particular type, $\hat{\theta}$, which is potentially not his true type. 
In this case, no matter
how much information he gathers against being of type $\hat{\theta}$, he will not update his beliefs. Any discrepancies between the 
observed outcomes and his believed type are incorporated using the Bayes rule to update their beliefs about $\omega$. 
This means the dogmatic agent will never learn their true type if they startu off as being misspecified.

@Heidhues2018 show that, under certain assumptions on the per-period utility,\footnote{The assumptions are that 
$u$ is twice continuously differentiable with: (i)$u_{ee}<0$ and $u_e(\underline{e} \theta, \omega)>0>u_e(\Bar{e}, \theta, \omega)$, 
(ii) $u_{\theta}, u_{\omega}>0$ and (iii) $u_{e\theta}<0$ and $u_{e\omega}>0$. The direction of the derivatives is a normalization
and the results would hold even when the signs are reversed.} 
a dogmatic modeler will inevitably fall into a self-defeating equilibrium. The equilibrium will be such that
the outcomes they observe reinforce their belief on $\omega$ in such a way that as $t\to\infty$ the agent will be sure that
the state is some $\omega^{'}$ consistent with their believed type and the observed data. In other words, 
they will be in a self-defeating equilibrium whith a stable belief that places probability one on the incorrect
parameters $(\hat{theta}, \omega_\infty)$

The mechanism by which the dogmatic agent falls into the self-defeating equilibrium is the following: Suppose the agent holds the 
misspecified belief that they are type $\hat{\theta}>\theta^*$. For any prior over $\omega$, the agent will be disappointed by the outcome.
He expected a gain of $h(\hat{\theta}, \mathbb{E}(\omega), a)$ but instead observes  $h(\hat{\theta}, \omega^*, a)$. 
There are two 
possible sources for the disappointment, the first is that the realized state is lower than the expected state. The second source is that the
agent is of type $\theta^*$ and therefore, for all possible states, his gain will be lower than what he expected. Because the agent is dogmatic,
he will not update his beliefs about $\theta$ and as a consequence will attribute the disappointment to the state being lower than expected.
He will continue to update in ths way until he converges to a belief about $\omega$ that is stable. Such a belief will 
explain the observed utility perfectly and allow the agent to rationalize his dogmatic belief about $\theta$. 
Under the assumptions of @Heidhues2018, there is a unique value of $\omega$ at which the belief is stable, I will refer to such value
as $\omega_\infty$.
This mechanism is further illustrated in Example 1.

**Example 1: ** 
_Set $A = \Omega$ and $H = [0, \infty)$ and consider a student with intrinsic ability $\theta^*\geq 0$ who faces a grading procedure 
$\omega^*$ that is unknown to them. However, 
they know that a higher $\omega^*$ is more likely to yield a higher grade. In particular, assume the grade is given by 
$(\theta^*+a)\omega^*$.

The student must choose an effort level $a$, which determines their grade. For whatever the chosen effort level, 
the agent must pay a cost $c(a) = \frac{1}{2}a^2$. And he repeats this process for infinitely many periods. 
Assume also that the student's prior is such that $mathbb{E}[\omega]= \omega^*$ and he is dogmatic about being of 
type $\hat{\theta}>\theta^*. \footnote{The example is illustrated for an overconfident agent but the results are symmetric for a digmatic agent who initially 
places probability one on some $\tilde{\theta}<\theta^*$.}
Therefore, the student's payoff in period $t$ is given by

\begin{equation}
u_t(a_t; \theta^*, \omega^*) = (\theta^*+a_t)\omega^* - \frac{1}{2}a^2 + \varepsilon_t
\end{equation}

Under this specification, the myopic optimal effort level is $a_t^* = \omega^*$. Nonetheless, because the agent does not know $\omega^*$,
he will choose $a_t = \mathbb{E}_t(\omega)$ where the expectation is taken with respect to the agent's belief at the beginning of
period $t$. If he does not revise his effort choice for $k$ periods, he will receive 
an average utility of 
$(\theta^*+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}$ but he was expecting an average utility of 
$(\hat{\theta}\theta+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}$. In response, he will apply bayes rule to update his beliefs about
$\omega$ to get the posterior belief with $\mathbb_{t+k}[\omega] = \frac{(\theta^* + \omega^*)\omega^*}{(\hat{\theta} + \omega^*)}$
which is lower than the initial belief. This will cause the agent to choose a lower effort at $t+k$. As a result, he will again
receive an average utlitity that is lower than what he expected which will cause his belief to drift further down. This process will
continue until the average utility equals his expected utility under the dogmatic belief that assigns probability 1 to $\hat{\theta}$. 
At that point, the stundent will have reached a self-defeating equilibrium and he will continue to choose sub-optimal effort forever._

Although the model of a dogmatic modeler is able to rationalize the prevalence of overconfident (underconfident) beliefs, the
assumption that the agent has a degenerate belief and no mechanism through which he can update such belief is very restrictive. An
alternative approach is proposed by @Ba2023. She proposes an extension of the dogmatic agent who is able to jump from one dogmatic 
belief to another. By doing so, the agent might end up being dogmatic and correctly speciefied.

@@

An agent is a \emph{switcher} if they behave as a dogmatic, but is willing to entertain the possibility that they are of a different type.
In particular, when they start off as a misspecified dogmatic, they are willing to switch to a different dogmatic belief if the data is
convincing enough. Notice that their beliefs are still degenerate and assign probability one to a particular type, and zero to 
all other types. This means that a bayesian update on $\theta$ would cause no change in their beliefs.
However, they are willing to entertain two such beliefs and have a mechanism by which they decide which belief to adopt at any period $t$.

@@

** Example 1 (continued)** 
_ Take the same student from the example and now add the possibility of switching beliefs. _

# Seting 2

## The Bayesian

Bayesian agents always choose the effort level that maximizes their flow payoff by taking expectations over
their prior beliefs about $\theta$ and $\omega$. Since agents are myopic, even though all the parameters could be identified with enough variation in choices,

a fully Bayesian agent might not learn their true type. This happens because they do not internalize the tradeoff between flow payoff
and learning and thus might not experiment enough to learn their type. An alternative to this approach is given 
by @Hestermann2021 and is discussed with the results.

## The Biased Agent
An agent that updates their beliefs with self-attribution bias will update their beliefs about the state $\omega$ and their type $\theta$
by over-attributing successes to a high value of $\theta$ and under-estimating the role of higher $\omega$. Similarly,
they will attribute failure to a low state more than an unbiased agent would. 
To model the self-serving attribution bias, I take the approach of @benjamin2019, where a generalization of the Bayes rule above gives the update.

# A Unifying Framework
The agent can be of one of 3 types: $\theta \in \{\theta_L, \theta_M, \theta_H\}$  with $\theta_H > \theta_M > \theta_L$.
They face an unknown exogenous success rate
$\omega \in \{\omega_L, \omega_M, \omega_H\}$ with $\omega_H>\omega_M>\omega_L$. Each of the values of $\omega$ is realized with equal probability. 
The agent knows the distribution of $\omega$ but not its realized value. Let the true type be $\theta^*$ 
and let the true state be $\omega^*$.

@@

## The Dogmatic Modeler

A dogmatic modeler updates their beliefs about omega by using bayes rule. For the example, that means he ends up with the following
posterior odds ratios:

### 2023-09-05 Paper/draft.tex
editing outline

\title{Learning from Data Through Models}
\author{Jimena Galindo}
\date{September 04, 2023}

@@

\hypertarget{framework}{%
\section{Framework}\label{framework}}

REWRITE IN GENERAL TERMS WITH THE PROBABILITY BEING GIVEN BY A FUNCTION
THAT SATIEFIES ALL THE PROPERTIES IN BA SO THAT I CAN REFER TO HER
RESULTS. THEN EXPAND ON THE EXAMPLE FOR THE LAB AND SAY WHY IT CAPTURES
ALL THE FORCES WE WANT.

@@

Bayesian agents always choose the effort level that maximizes their flow
payoff by taking expectations over their prior beliefs about \(\theta\)
and \(\omega\). Since agents are myopic, even though all the parameters
could be identified with enough variation in choices, a fully Bayesian
agent might not learn their true type. This happens because they do not
internalize the tradeoff between flow payoff and learning and thus might
not experiment enough to learn their type. An alternative to this
approach is given by \citet{Hestermann2021} and is discussed with the
results.

\hypertarget{a-biased-agent}{%
\subsection{A Biased Agent}\label{a-biased-agent}}

@@

\hypertarget{the-dogmatic-agent}{%
\subsection{The Dogmatic Agent}\label{the-dogmatic-agent}}

A dogmatic agent does not update their beliefs about \(\theta\);
instead, they hold a degenerate belief that places probability one on
being a particular type, \(\hat{\theta}\). For instance, the dogmatic
agent might be of type \(\hat{\theta} = \theta_M\) but holds a belief
that places probability one on being of type \(\theta_H\). In this case,
no matter how much information he gathers against being of type
\(\theta_H\), he will not update his beliefs. Any discrepancies between
the observed outcomes are incorporated using the Bayes rule to update
their beliefs about \(\omega\). This results in the following posterior
odds:

A crucial difference between the dogmatic agent and the Bayesian agent
is that the dogmatic agent does not aggregate across types when updating
their beliefs about \(\omega\). This means the dogmatic agent will never
learn their true type if they are misspecified.

\citet{Heidhues2018} show that in a setting such as ours, a dogmatic
modeler will inevitably fall into a self-confirming equilibrium where
the outcomes they observe reinforce their belief on \(\omega\) in such a
way that as \(t\to\infty\) the agent will be certain that the state is
some \(\omega^*\) consistent with their believed type and the observed
data.

EXPLAIN WHY THEY ARE ALWAYS DISAPPOINTED AND UPDATE DOWNWARDS ON OMEGA
UNTIL THEY FALL INTO THE TRAP.

\[\frac{p[h^t|\tilde{\theta}}{p[h^t|\hat{\theta}]} > \alpha\]

@@

The agent can be of one of 3 types:
\(\theta \in \{\theta_L, \theta_M, \theta_H\}\) with
\(\theta_H > \theta_M > \theta_L\). They face an unknown exogenous
success rate \(\omega \in \{\omega_L, \omega_M, \omega_H\}\) with
\(\omega_H>\omega_M>\omega_L\). Each of the values of \(\omega\) is
realized with equal probability. The agent knows the distribution of
\(\omega\) but not its realized value. Let the true type be \(\theta^*\)
and the true state be \(\omega^*\).

@@

In the data-generating process described above, there are six such
equilibria. For example, if the agent is of type \(\theta_M\) but
mistakenly believes that he is of type \(\hat{\theta}=\theta_H\) and the
and \(\omega^* = \omega_M\), when the effort chosen is \(e_L\), the
agent will observe a success with 45\% chance. Because the agent
dogmatically believes that their type is high, they will erroneously
conclude that the rate is \(\omega_L\). Under this belief, the optimal
action is \(e_L\) which will continue to generate successes with \(45%
\) probability, further reinforcing the incorrect belief. If the agent
were to learn the true type and rate, they would instead choose \(e_M\),
yielding a success with \(65%
\) probability and strictly higher expected payoffs.

@@

>>>>

\title{Learning from Data Through Models}
\author{Jimena Galindo}
\date{September 05, 2023}

@@

\hypertarget{related-literature}{%
\section{Related Literature}\label{related-literature}}

\hypertarget{framework-1}{%
\section{Framework 1}\label{framework-1}}

An agent is of type \(\theta \in \Theta\) and faces an unknown exogenous
state \(\omega\) drawn from some density \(f\) over \(\Omega\). The
agent knows the distribution of \(\omega\) but not its realized value.
His prior belief about his type is \(p_0(\theta)\) and his belief about
the state state \(p_0(\omega)\) coincides with the true distribution
\(f\). Let the agent's type be \(\theta^*\) and the realized state be
\(\omega^*\).

An agent has a \emph{misspecified} belief if the prior assigns
probability zero to their true type. Furthermore, the agent is
\emph{dogmatic} if he holds a degenerate belief that places probability
one on being a particular type, \(\hat{\theta}\). A dogmatic agent can
be dogmatic and misspecified. That means that
\(\hat{\theta} \neq \theta^*\) and \(p_0(\hat{\theta}) = 1\).

The agent chooses an action \(a\in A\) and observes a noisy outcome
\(h\in H\). The outcome is a function of the agent's type, the state,
and the action. In particular
\(h = h(\theta^*, \omega^*, a) + \varepsilon\) with \(h(\cdot)\)
increasing in both \(\theta^*\) and \(\omega^*\); and such that
conditional on a pair of parameters \((\theta, \omega)\), there is a
unique optimal action. \(\varepsilon\sim N(0, \sigma)\) is noise in the
output.

After observing the outcome, the agent updates his beliefs about
\(\theta\) and \(\omega\) using some algorithm and moves on to the next
period. He repeats this process infinitely many times. I make the
simplifying assumption that the agent is myopic and chooses the action
that maximizes the payoff in each period. This assumption is crucial for
the results as the agent will not trade off the flow payoff for learning
the truth as a forward-looking agent would.

Within this framework, I consider two nested theories of belief
updating. The first one is a dogmatic modeler from \citet{Heidhues2018}.
The second one is a switcher, as in \citet{Ba2023}. The dogmatic modeler
can be seen as a switcher with an infinitely sticky initial belief and
this is the sense in which the two are nested.

A key notion in this setting is that of a self-defeating
equilibrium\footnote{This notion is an adaptation of the Berk-Nash equilibrium in 
@Esponda2016}. A \emph{self-defeating equilibrium} is a belief and
action pair such that the agent's belief about their type is
misspecified and the outcome generated by the action is consistent with
the misspecified belief. The average outcome under the true type and the
true state equals the average output the agent expects under the
misspecified belief. In addition, the agent's belief is said to be
\emph{stable} when this happens.

\hypertarget{the-dogmatic-modeler}{%
\subsection{The Dogmatic Modeler}\label{the-dogmatic-modeler}}

A dogmatic agent does not update their beliefs about \(\theta\); instead
he holds a degenerate belief that places probability one on being a
particular type, \(\hat{\theta}\), which is potentially not his true
type. In this case, no matter how much information he gathers against
being of type \(\hat{\theta}\), he will not update his beliefs. Any
discrepancies between the observed outcomes and his believed type are
incorporated using the Bayes rule to update their beliefs about
\(\omega\).

A key difference between the dogmatic agent and a fully Bayesian agent
is that the dogmatic agent does not aggregate across types when updating
their beliefs about \(\omega\). This means the dogmatic agent will never
learn their true type if they are misspecified.

\citet{Heidhues2018} show that, under certain assumptions on the
per-period utility,\footnote{The assumptions are that 
$u$ is twice continuously differentiable with: (i)$u_{ee}<0$ and $u_e(\underline{e} \theta, \omega)>0>u_e(\Bar{e}, \theta, \omega)$, 
(ii) $u_{\theta}, u_{\omega}>0$ and (iii) $u_{e\theta}<0$ and $u_{e\omega}>0$. The direction of the derivatives is a normalization
and the results would hold even when the signs are reversed.} about the
outcome \(h\), a dogmatic modeler will inevitably fall into a
self-defeating equilibrium where the outcomes they observe reinforce
their belief on \(\omega\) in such a way that as \(t\to\infty\) the
agent will be sure that the state is some \(\omega^{'}\) consistent with
their believed type and the observed data.

The mechanism by which the dogmatic agent falls into the self-defeating
equilibrium is the following: Suppose the agent holds the misspecified
belief that they are type \(\hat{\theta}>\theta^*\). For any prior over
\(\omega\), the agent will be disappointed by the outcome; he expected a
gain of \(h(\hat{\theta}, \mathbb{E}(\omega), a)\) but instead observes
\(h(\hat{\theta}, \omega^*, a)\). There are two possible sources for the
disappointment. The first is that the realized state is lower than the
expected state. The second is that the agent is of type \(\theta^*\) and
thus for all possible states, his gain will be lower than what he
expected. Because the agent is dogmatic, he will not update his beliefs
about \(\theta\) and thus will attribute the disappointment to the state
being lower than expected. \citet{Heidhues2018} give sufficient
conditions for the agent to fall into a self-confirming equilibrium
where the agent will continue to choose an action that reinforces an
incorrect belief about the state. This force is illustrated in Example
1.

\textbf{Example 1: } Set \(A = \Omega\) and \(H = [0, \infty)\) and
consider a student with intrinsic ability \(\theta^*\geq 0\) who faces a
grading procedure \(\omega^*\) that is unknown to them. However, they
know that a higher \(\omega^*\) is more likely to yield a higher grade.
In particular, assume the grade is given by \((\theta^*+a)\omega^*\).

The student must choose an effort level \(a\), which determines their
grade. For whatever the chosen effort level, the agent must pay a cost
\(c(a) = \frac{1}{2}a^2\). And he repeats this process for infinitely
many periods. Assume also that the student's prior is such that
\(mathbb{E}[\omega]= \omega^*\). Therefore, the student's payoff in
period \(t\) is given by

\begin{equation}
u_t(a_t; \theta^*, \omega^*) = (\theta^*+a_t)\omega^* - \frac{1}{2}a^2 + \varepsilon_t
\end{equation}

Under this specification, the myopic optimal effort level is
\(a_t^* = \omega^*\). Nonetheless, because the agent does not know
\(\omega^*\), he will choose \(a_t = \mathbb{E}_t(\omega)\) where the
expectation is taken with respect to the agent's belief at the beginning
of period \(t\). If he does not revise his effort choice for multiple
periods, he will observe an average grade of
\((\theta^*+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}\) but he was expecting
an average grade of
\((\hat{\theta}\theta+a_t^*)\omega^* - \frac{1}{2}a_t^{*2}\)

@@

Bayesian agents always choose the effort level that maximizes their flow
payoff by taking expectations over their prior beliefs about \(\theta\)
and \(\omega\). Since agents are myopic, even though all the parameters
could be identified with enough variation in choices,

\[\frac{p[h^t|\tilde{\theta}]}{p[h^t|\hat{\theta}]} > \alpha\]

\hypertarget{the-bayesian-agent}{%
\subsection{The Bayesian Agent}\label{the-bayesian-agent}}

a fully Bayesian agent might not learn their true type. This happens
because they do not internalize the tradeoff between flow payoff and
learning and thus might not experiment enough to learn their type. An
alternative to this approach is given by \citet{Hestermann2021} and is
discussed with the results.

\hypertarget{the-biased-agent}{%
\subsection{The Biased Agent}\label{the-biased-agent}}

@@

@@

The agent can be of one of 3 types:
\(\theta \in \{\theta_L, \theta_M, \theta_H\}\) with
\(\theta_H > \theta_M > \theta_L\). They face an unknown exogenous
success rate \(\omega \in \{\omega_L, \omega_M, \omega_H\}\) with
\(\omega_H>\omega_M>\omega_L\). Each of the values of \(\omega\) is
realized with equal probability. The agent knows the distribution of
\(\omega\) but not its realized value. Let the true type be \(\theta^*\)
and let the true state be \(\omega^*\).

@@

In the data-generating process described above, there are five such
equilibria. For example, if the agent is of type \(\theta_M\) but
mistakenly believes that he is of type \(\hat{\theta}=\theta_H\) and the
and \(\omega^* = \omega_M\), when the effort chosen is \(e_L\), the
agent will observe a success with 45\% chance. Because the agent
dogmatically believes that their type is high, they will erroneously
conclude that the rate is \(\omega_L\). Under this belief, the optimal
action is \(e_L\) which will continue to generate successes with \(45%
\) probability, further reinforcing the incorrect belief. If the agent
were to learn the true type and rate, they would instead choose \(e_M\),
yielding a success with \(65%
\) probability and strictly higher expected payoffs.

@@

\hypertarget{the-dogmatic-modeler-1}{%
\subsection{The Dogmatic Modeler}\label{the-dogmatic-modeler-1}}

A dogmatic modeler updates their beliefs about omega by using bayes
rule. For the example, that means he ends up with the following
posterior odds ratios:
